---
output:
  pdf_document: default
  html_document: default
---
# Frequency of french names by department and year


The aim of the activity is to develop a methodology to answer a specific question on a given dataset. 

The dataset is the set of Firstname given in France on a large period of time. 
[given names data set of INSEE](https://www.insee.fr/fr/statistiques/fichier/2540004/dpt2017_txt.zip), we choose this dataset because it is sufficiently large, you can't do the analysis by hand, the structure is simple

## Download Raw Data from the website
```{r}
file = "dpt2019_txt.zip"
if(!file.exists(file)){
  download.file("https://www.insee.fr/fr/statistiques/fichier/2540004/dpt2019_csv.zip",
	destfile=file)
}
unzip(file)
```

## Build the Dataframe from file. We see that the data has rows with unexpected values.

```{r}
library(tidyverse)
library(ggplot2)
FirstNames <- read_delim("dpt2019.csv",delim=";");
```

## Show structure of the loaded data and clean from invalid values

```{r}
str(FirstNames)

```
 Clean data from wrong types

```{r}
unique(select(problems(FirstNames), col, actual))
```

Removing data where the year and name are unknown

```{r}
FirstNamesClean <- filter(FirstNames, annais!="XXXX" & preusuel!="_PRENOMS_RARES")
```


## Analysis of a first name occurence over time. Comparison of frequency of several firstnames


Here we are showing frequency along the years of 3 different male names that start with the letter H. We can see that names Hubert and Henri were popular around the beginning of the century, Hugues around the middle and that the name Hugo has become popular around the end of the century.

```{r}
Hubert <- summarise(group_by(select(filter(FirstNamesClean, preusuel=="HUBERT"), annais, nombre), annais), nombre=sum(nombre))
ggplot(data=Hubert, aes(x=annais, y=nombre))+geom_bar(stat="identity")
```
```{r}
Hugo <- summarise(group_by(select(filter(FirstNamesClean, preusuel=="HUGO"), annais, nombre), annais), nombre=sum(nombre))
ggplot(data=Hugo, aes(x=annais, y=nombre))+geom_bar(stat="identity")
```


```{r}
Henri <- summarise(group_by(select(filter(FirstNamesClean, preusuel=="HENRI"), annais, nombre), annais), nombre=sum(nombre))
ggplot(data=Henri, aes(x=annais, y=nombre))+geom_bar(stat="identity")
```
```{r}
Hugues <- summarise(group_by(select(filter(FirstNamesClean, preusuel=="HUGUES"), annais, nombre), annais), nombre=sum(nombre))
ggplot(data=Hugues, aes(x=annais, y=nombre))+geom_bar(stat="identity")
```


## Establish by gender the most given firstname by year. Analyse the evolution of the most frequent firstname.

```{r}

HighestOccurence <- unique(filter(group_by(summarise(group_by(FirstNamesClean, annais, preusuel, sexe), sum_nombre = sum(nombre), sexe), annais), sum_nombre == max(sum_nombre)))
  
  #filter(), n == max(sum_nombre))

HighestOccurenceWomen <- filter(HighestOccurence, sexe==2)
HighestOccurenceMen <- filter(HighestOccurence, sexe==1)

```


### Frequency of most frequent male names by year

```{r}
ggplot(data = HighestOccurenceMen, aes(x=annais, y=sum_nombre, color = preusuel))+geom_point()

```

# Frequency of most frequent female names by year

```{r}
ggplot(data = HighestOccurenceWomen, aes(x=annais, y=sum_nombre, color = preusuel))+geom_point()
```